---
title: "materialstudy"
---


This website is a student project that explores the possibilities of graphing using ggplot2 in rstudio. The data used was sourced from the "Dutch Textile Trade" digital project. You can visit the project [here](https://dutchtextiletrade.org/).

```{r, include=FALSE}
read.csv("full_textile_data.csv")
dtd <- read.csv("full_textile_data.csv")
```

The data contained in the Data Set tracks the trade of textiles from 1699 to 1724. It records variables such as quantities, dates, total cost, and textile type. 

For this project I will be plotting total cost of textiles traded over time. The figure below plots this using a default geom_point graph. This is a good start, however the format can be improved.

```{r}
library(ggplot2)
ggplot(data = dtd,
       mapping = aes(x = dest_yr,
           y = total_value)) +
        geom_point()
```

In this graph the visualization is improved with the usage of a bar graph. However, there are issues with the scale and the readability. Ggplot cannot successfully read the "total value" column in the original excel sheet.

```{r}
ggplot(data = dtd,
       mapping = aes(x = dest_yr,
             y = total_value)) +
             geom_col(fill="steelblue")
```

Defining the x axis as continuous modifies the data to make it legible to ggplot. However, this graph is still not communicating the data in a useful way.


```{r}
ggplot(data = dtd,
        mapping = aes(y = factor(dest_yr),
              x = total_value)) +
              geom_col(fill = "steelblue") +
   scale_x_continuous(labels = scales::label_comma())
```


