---
title: "material study"
---


This website is a student project that explores the possibilities of graphing using ggplot2 in rstudio. The data used was sourced from the "Dutch Textile Trade" digital project. You can visit the project [here](https://dutchtextiletrade.org/).

```{r, include=FALSE}
read.csv("full_textile_data.csv")
dtd <- read.csv("full_textile_data.csv")
```

The data contained in the Data Set tracks the trade of textiles from 1699 to 1724. It records variables such as quantities, dates, total cost, and textile type. 

For this project I will be plotting total cost of textiles traded over time. The figure below plots this using a default geom_point graph. This is a good start, however the format can be improved.

```{r}
library(ggplot2)
ggplot(data = dtd,
       mapping = aes(x = dest_yr,
           y = total_value)) +
        geom_point()
```

Line graphs can be effective at showing change over time. The figure below shows a graph using geom_line rather than the geom_point function. However, this clearly is still not useful for visualizing the data. The line is cramped at the beginning, and the actual changes in the line doesn't make sense.

```{r}
 dtd |>
   ggplot(mapping = aes(y = dest_yr,
                        x = total_value)) +
                   geom_line()
```


In this figure the visualization is improved with the usage of a bar graph. However, there are issues with the scale and the readability. Ggplot cannot successfully read the "total value" column in the original excel sheet, and the bars also feel cramped and dificult to read.

```{r}
ggplot(data = dtd,
       mapping = aes(x = dest_yr,
             y = total_value)) +
             geom_col(fill="steelblue")
```

Defining the x axis as continuous modifies the data to make it legible to ggplot. However, this graph is still not communicating the data in a useful way.


```{r}
ggplot(data = dtd,
        mapping = aes(y = factor(dest_yr),
              x = total_value)) +
              geom_col(fill = "steelblue") +
   scale_x_continuous(labels = scales::label_comma())
```


This figure finally presents legible information. The years are defined and it is clear what kinds of changes are occurring over time. In order to accomplish this I used the filter function to display total cost over time for one specific textile called "adathaies".



```{r}
dtd |>
 filter(textile_name == "adathaies") |>
   ggplot(mapping = aes(y = factor(dest_yr),
                        x = total_value)) +
   geom_col(fill = "steelblue") +
   scale_x_continuous(labels = scales::label_comma())
```

Something important to note is that while the graph above is effective at visualizing this specific data set, this does not indicate that there is one specific way that data should be graphed or plotted. By documenting my process for this project, I hope to showcase that graphing is not a one-step activity, but rather a process that requires experimentation and trial and error.

